---
title: "Data Science Capstone - Final Project"
author: "Famibelle Médhi"
date: "Saturday, April 25, 2015"
output: ioslides_presentation
---

## Introduction
The project aim is to build a coherent text prediction app hosted on shiny. It aims at predicting from an input phrase, what would be the next word in the sentence.

The application is the final deliverable of the **Coursera Data Science Specialization**. 
All the calculation were made on a basic 64 bits laptop with 4Go of RAM and 2.70GHz

## Files description

The Corpus is made of three text files extracted from https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip

    - en_US.news.txt file size is 205 811 889 bytes

    - en_US.blogs.txt file size is 210 160 014 bytes

    - en_US.twitter.txt file size is 167 105 338 bytes
    

## Data
First I created a Corpus and extract from the Corpus n-Grams, (i.e. unigrams, bigrams, trigrams, 4-grams, 5-grams). Then the 

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

