---
title: "Data Science Capstone - Milestone Report"
author: "Famibelle Médhi"
date: "Wednesday, March 25, 2015"
output: html_document
---


```{r setOptions, message=FALSE, echo=FALSE, warning=FALSE}
# Load all the needed library
library(RWeka)
library(stringi)
library(tm)
library(SnowballC)
library(reshape2)
library(openNLP)
library(RWeka)
library(googleVis)
op <- options(gvis.plot.tag='chart')
library(qdap)
library(wordcloud)
```

## load the data
Get all the data are located from https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip
All downloaded data are in zip format. From the downloaded data we are going to work on the following files 

1. en_US.news.txt file size is `r format(file.info("en_US.news.txt")$size, digits=9, decimal.mark=".", big.mark=" ")` bytes 

1. en_US.blogs.txt file size is  `r format(file.info("en_US.blogs.txt")$size, digits=9, decimal.mark=".", big.mark=" ")` bytes

1. en_US.twitter.txt file size is `r format(file.info("en_US.twitter.txt")$size, digits=9, decimal.mark=".", big.mark=" ")` bytes


```{r loadData, echo=FALSE, warning=FALSE}
# Loading the data
con <- file("en_US.twitter.txt", "r")
twitter <- readLines(con)
close(con)

con <- file("en_US.blogs.txt", "r")
blogs <- readLines(con)
close(con)

con <- file("en_US.news.txt", "r")
news <- readLines(con)
close(con)
rm(con)

twitter.vector  <- VectorSource(twitter)
blogs.vector    <- VectorSource(blogs)
news.vector     <- VectorSource(news)
```


## Very basic analysis of the three files

file *en_US.twitter.txt* contains `r format(twitter.vector$length, digits=9, decimal.mark=".", big.mark=" ")` lines, `r format(length(unlist(stri_extract_all_words(twitter))), digits=9, decimal.mark=".", big.mark=" ")` words and `r format(sum(nchar(twitter)), digits=9, decimal.mark=".", big.mark=" ")` characters

file *en_US.blogs.txt* contains `r format(blogs.vector$length, digits=9, decimal.mark=".", big.mark=" ")` lines, `r format(length(unlist(stri_extract_all_words(blogs))), digits=9, decimal.mark=".", big.mark=" ")` words and `r format(sum(nchar(blogs)), digits=9, decimal.mark=".", big.mark=" ")` characters

file *en_US.news.txt* contains `r format(news.vector$length, digits=9, decimal.mark=".", big.mark=" ")` lines, `r format(length(unlist(stri_extract_all_words(news))), digits=9, decimal.mark=".", big.mark=" ")` words and `r format(sum(nchar(twitter)), digits=9, decimal.mark=".", big.mark=" ")` characters


```{r echo=FALSE}
summary <- as.data.frame(c("en_US.news.txt", "en_US.blogs.txt", "en_US.twitter.txt"))

summary$filesize <- c(
    file.info("en_US.news.txt")$size,
    file.info("en_US.blogs.txt")$size,
    file.info("en_US.twitter.txt")$size
)
summary$lines <- c(
    news.vector$length,
    blogs.vector$length,
    twitter.vector$length
    )

summary$words <- c(
    length(unlist(stri_extract_all_words(news))),
    length(unlist(stri_extract_all_words(blogs))),
    length(unlist(stri_extract_all_words(twitter)))
    )

summary$characters <- c(
    sum(nchar(news)),
    sum(nchar(blogs)),
    sum(nchar(twitter))
    )      
           
summary$min_character_per_line <- c(
    min(nchar(news)),
    min(nchar(blogs)),
    min(nchar(twitter))
    )

summary$max_character_per_line <- c(
    max(nchar(news)),
    max(nchar(blogs)),
    max(nchar(twitter))
    )

summary$mean_character_per_line <- c(
    mean(nchar(news)),
    mean(nchar(blogs)),
    mean(nchar(twitter))
    )

colnames(summary) <- c(
    "file names",
    "file size (in bytes)",
    "Number of lines",
    "Number of Words",
    "Number of characters",
    "Minimum number of characters per line",
    "Maximum number of characters per line",
    "Mean number of characters per line"
    )
summary_plot <- gvisTable(
    summary,
    formats=list(
        "file size (in bytes)" = "#,###",
        "Number of lines" = "#,###",
        "Number of Words" = "#,###",
        "Number of characters" = "#,###",
        "Minimum number of characters per line" = "#,###",
        "Maximum number of characters per line" = "#,###",
        "Mean number of characters per line" = "#,###"
        )
    )
```

```{r results="asis"}
print(summary_plot, "chart")
```

We can see from the summary table that files are huge and beyond commun computers processing power. We will need to sample the files to draw first conclusions.
The **en_US.twitter.txt** file has the largest number of lines
the **en_US.blogs.txt** file has largest number of words and the largest number of characters
the **en_US.blogs.txt** has by far the largest number of character per line.

## Building the sample dataset by sampling the files
```{r echo=FALSE}
sample.size <- 1000
```

Creating corpus directly from the files needs is too CPU consuming and would take to much time. So I decided to sample **`r sample.size` samples** of the files content in order to draw the first conclusions. 

```{r}
twitter.sample  <- sample(twitter, size = sample.size)
blogs.sample    <- sample(blogs, size = sample.size)
news.sample     <- sample(x = news, size = sample.size)
```

## Create a corpus 

Create a corpus and remove punctuation, whitespace, numbers, transform and the text to lower case and stem document. 

```{r CorpusCreation}
# Create a corpus
Create_Corpus <- function(sample) {
    Corpus <- VCorpus(VectorSource(sample))
    Corpus <- tm_map(Corpus, removePunctuation)
    Corpus <- tm_map(Corpus, stripWhitespace)
    Corpus <- tm_map(Corpus, removeNumbers)
    Corpus <- tm_map(Corpus, content_transformer(tolower))
#     Corpus <- tm_map(Corpus, removeWords, Profanities)
    Corpus <- tm_map(Corpus, stemDocument) # Stem document
    return(Corpus)
}

twitter.corpus  <- Create_Corpus(twitter.sample)
blogs.corpus    <- Create_Corpus(blogs.sample)
news.corpus     <- Create_Corpus(news.sample)
```

## frequencies of the 1-grams, 2-grams and 3-grams in the sample dataset
```{r Tokenization}
One_Gram_Tokenizer <- function(character_vector) { NGramTokenizer(character_vector, Weka_control(min = 1, max = 1))}
Two_Gram_Tokenizer <- function(character_vector) { NGramTokenizer(character_vector, Weka_control(min = 2, max = 2))}
Tri_Gram_Tokenizer <- function(character_vector) { NGramTokenizer(character_vector, Weka_control(min = 3, max = 3))}

twitter.One_Gram <- TermDocumentMatrix(twitter.corpus, control = list( tokenize = One_Gram_Tokenizer))
twitter.Two_Gram <- TermDocumentMatrix(twitter.corpus, control = list( tokenize = Two_Gram_Tokenizer))
twitter.Tri_Gram <- TermDocumentMatrix(twitter.corpus, control = list( tokenize = Tri_Gram_Tokenizer))

blogs.One_Gram <- TermDocumentMatrix(blogs.corpus, control = list( tokenize = One_Gram_Tokenizer))
blogs.Two_Gram <- TermDocumentMatrix(blogs.corpus, control = list( tokenize = Two_Gram_Tokenizer))
blogs.Tri_Gram <- TermDocumentMatrix(blogs.corpus, control = list( tokenize = Tri_Gram_Tokenizer))

news.One_Gram <- TermDocumentMatrix(news.corpus, control = list( tokenize = One_Gram_Tokenizer))
news.Two_Gram <- TermDocumentMatrix(news.corpus, control = list( tokenize = Two_Gram_Tokenizer))
news.Tri_Gram <- TermDocumentMatrix(news.corpus, control = list( tokenize = Tri_Gram_Tokenizer))
```

### *en_US.twitter.txt* dataset sample n-grams frequency analysis

```{r echo=FALSE}
top <- 50
twitter.One_Gram_Analysis <- sort(rowSums(as.matrix(twitter.One_Gram)), decreasing = TRUE)
twitter.One_Gram_Analysis <- as.data.frame(as.matrix(twitter.One_Gram_Analysis))
twitter.One_Gram_Analysis$TriGram <- rownames(twitter.One_Gram_Analysis)
twitter.One_Gram_Analysis <- twitter.One_Gram_Analysis[,c(2,1)]
colnames(twitter.One_Gram_Analysis) <- c("twitter.1-gram", "count")

One_Gram_Plot_twitter <- gvisColumnChart(
    head(twitter.One_Gram_Analysis, top),
    xvar = "twitter.1-gram",
    yvar = "count",
    options=list(
        width=900,height=400,
        title = paste("The top ", top ," 1-grams with Highest Frequency over n=", sample.size, " samples from the ", summary[3,1] ," dataset", sep="")
        )
    )

twitter.Two_Gram_Analysis <- sort(rowSums(as.matrix(twitter.Two_Gram)), decreasing = TRUE)
twitter.Two_Gram_Analysis <- as.data.frame(as.matrix(twitter.Two_Gram_Analysis))
twitter.Two_Gram_Analysis$TriGram <- rownames(twitter.Two_Gram_Analysis)
twitter.Two_Gram_Analysis <- twitter.Two_Gram_Analysis[,c(2,1)]
colnames(twitter.Two_Gram_Analysis) <- c("twitter.2-gram", "count")

Two_Gram_Plot_twitter <- gvisColumnChart(
    head(twitter.Two_Gram_Analysis, top),
    xvar = "twitter.2-gram",
    yvar = "count",
    options=list(
        width=900,height=400,
        title = paste("The top ", top, " 2-grams with Highest Frequency over n=", sample.size, " samples from the ", summary[3,1] ," dataset", sep="")
        )
    )

twitter.Tri_Gram_Analysis <- sort(rowSums(as.matrix(twitter.Tri_Gram)), decreasing = TRUE)
twitter.Tri_Gram_Analysis <- as.data.frame(as.matrix(twitter.Tri_Gram_Analysis))
twitter.Tri_Gram_Analysis$TriGram <- rownames(twitter.Tri_Gram_Analysis)
twitter.Tri_Gram_Analysis <- twitter.Tri_Gram_Analysis[,c(2,1)]
colnames(twitter.Tri_Gram_Analysis) <- c("twitter.3-gram", "count")

Tri_Gram_Plot_twitter <- gvisColumnChart(
    head(twitter.Tri_Gram_Analysis, top),
    xvar = "twitter.3-gram",
    yvar = "count",
    options=list(
        width=900,height=400,
        title = paste("The top ", top, " 3-grams with Highest Frequency over n=", sample.size, " samples from the ", summary[3,1] ," dataset", sep="")

        )
    )

n_Gram_Plot_twitter <- gvisMerge(One_Gram_Plot_twitter, Two_Gram_Plot_twitter, horizontal = FALSE)
n_Gram_Plot_twitter <- gvisMerge(n_Gram_Plot_twitter, Tri_Gram_Plot_twitter, horizontal = FALSE)
```

### plot the en_US.twitter.txt sample dataset n-grams
```{r OneGramTwitter, results="asis"}
print(n_Gram_Plot_twitter, "chart")
```

## en_US.blogs.txt dataset sample n-grams frequency analysis

```{r echo=FALSE}
blogs.One_Gram_Analysis <- sort(rowSums(as.matrix(blogs.One_Gram)), decreasing = TRUE)
blogs.One_Gram_Analysis <- as.data.frame(as.matrix(blogs.One_Gram_Analysis))
blogs.One_Gram_Analysis$TriGram <- rownames(blogs.One_Gram_Analysis)
blogs.One_Gram_Analysis <- blogs.One_Gram_Analysis[,c(2,1)]
colnames(blogs.One_Gram_Analysis) <- c("blogs.1-gram", "count")

One_Gram_Plot_blogs <- gvisColumnChart(
    head(blogs.One_Gram_Analysis, top),
    xvar = "blogs.1-gram",
    yvar = "count",
    options=list(
        width=900,height=400,
        title = paste("The top ", top, " 1-grams with Highest Frequency over n=", sample.size, " samples from the ", summary[2,1] ," dataset", sep="")

        )
    )

blogs.Two_Gram_Analysis <- sort(rowSums(as.matrix(blogs.Two_Gram)), decreasing = TRUE)
blogs.Two_Gram_Analysis <- as.data.frame(as.matrix(blogs.Two_Gram_Analysis))
blogs.Two_Gram_Analysis$TriGram <- rownames(blogs.Two_Gram_Analysis)
blogs.Two_Gram_Analysis <- blogs.Two_Gram_Analysis[,c(2,1)]
colnames(blogs.Two_Gram_Analysis) <- c("blogs.2-gram", "count")

Two_Gram_Plot_blogs <- gvisColumnChart(
    head(blogs.Two_Gram_Analysis, top),
    xvar = "blogs.2-gram",
    yvar = "count",
    options=list(
        width=900,height=400,
        title = paste("The top ", top, " 2-grams with Highest Frequency over n=", sample.size, " samples from the ", summary[2,1] ," dataset", sep="")
        )
    )

blogs.Tri_Gram_Analysis <- sort(rowSums(as.matrix(blogs.Tri_Gram)), decreasing = TRUE)
blogs.Tri_Gram_Analysis <- as.data.frame(as.matrix(blogs.Tri_Gram_Analysis))
blogs.Tri_Gram_Analysis$TriGram <- rownames(blogs.Tri_Gram_Analysis)
blogs.Tri_Gram_Analysis <- blogs.Tri_Gram_Analysis[,c(2,1)]
colnames(blogs.Tri_Gram_Analysis) <- c("blogs.3-gram", "count")

Tri_Gram_Plot_blogs <- gvisColumnChart(
    head(blogs.Tri_Gram_Analysis, top),
    xvar = "blogs.3-gram",
    yvar = "count",
    options=list(
        width=900,height=400,
        title = paste("The top ", top, " 3-grams with Highest Frequency over n=", sample.size, " samples from the ", summary[2,1] ," dataset", sep="")
        )
    )
n_Gram_Plot_blogs <- gvisMerge(One_Gram_Plot_blogs, Two_Gram_Plot_blogs, horizontal = FALSE)
n_Gram_Plot_blogs <- gvisMerge(n_Gram_Plot_blogs, Tri_Gram_Plot_blogs, horizontal = FALSE)
```

### plot the *en_US.blogs.txt* sample dataset n-grams
```{r OneGramblogs, results="asis"}
print(n_Gram_Plot_blogs, "chart")
```

## *en_US.news.txt* dataset sample n-grams frequency analysis

```{r echo=FALSE}
news.One_Gram_Analysis <- sort(rowSums(as.matrix(news.One_Gram)), decreasing = TRUE)
news.One_Gram_Analysis <- as.data.frame(as.matrix(news.One_Gram_Analysis))
news.One_Gram_Analysis$TriGram <- rownames(news.One_Gram_Analysis)
news.One_Gram_Analysis <- news.One_Gram_Analysis[,c(2,1)]
colnames(news.One_Gram_Analysis) <- c("news.1-gram", "count")

One_Gram_Plot_news <- gvisColumnChart(
    head(news.One_Gram_Analysis, top),
    xvar = "news.1-gram",
    yvar = "count",
    options=list(
        width=900,height=400,
        title = paste("The top ", top, " 3-grams with Highest Frequency over n=", sample.size, " samples from the ", summary[1,1] ," dataset", sep="")
        )
    )

news.Two_Gram_Analysis <- sort(rowSums(as.matrix(news.Two_Gram)), decreasing = TRUE)
news.Two_Gram_Analysis <- as.data.frame(as.matrix(news.Two_Gram_Analysis))
news.Two_Gram_Analysis$TriGram <- rownames(news.Two_Gram_Analysis)
news.Two_Gram_Analysis <- news.Two_Gram_Analysis[,c(2,1)]
colnames(news.Two_Gram_Analysis) <- c("news.2-gram", "count")

Two_Gram_Plot_news <- gvisColumnChart(
    head(news.Two_Gram_Analysis, top),
    xvar = "news.2-gram",
    yvar = "count",
    options=list(
        width=900,height=400,
        title = paste("The top ", top, " 2-grams with Highest Frequency over n=", sample.size, " samples from the ", summary[1,1] ," dataset", sep="")        
        )
    )

news.Tri_Gram_Analysis <- sort(rowSums(as.matrix(news.Tri_Gram)), decreasing = TRUE)
news.Tri_Gram_Analysis <- as.data.frame(as.matrix(news.Tri_Gram_Analysis))
news.Tri_Gram_Analysis$TriGram <- rownames(news.Tri_Gram_Analysis)
news.Tri_Gram_Analysis <- news.Tri_Gram_Analysis[,c(2,1)]
colnames(news.Tri_Gram_Analysis) <- c("news.3-gram", "count")

Tri_Gram_Plot_news <- gvisColumnChart(
    head(news.Tri_Gram_Analysis, top),
    xvar = "news.3-gram",
    yvar = "count",
    options=list(
        width=900,height=400, 
        title = paste("The top ", top, " 3-grams with Highest Frequency over n=", sample.size, " samples from the ", summary[1,1] ," dataset", sep="")
        )
    )

n_Gram_Plot_news <- gvisMerge(One_Gram_Plot_news, Two_Gram_Plot_news, horizontal = FALSE)
n_Gram_Plot_news <- gvisMerge(n_Gram_Plot_news, Tri_Gram_Plot_news, horizontal = FALSE)
```

### plot the *en_US.news.txt* sample dataset n-grams
```{r OneGramnews, results="asis"}
print(n_Gram_Plot_news, "chart")
```